\documentclass[9.5pt, a4paper]{article}

% ── PACKAGES ──────────────────────────────────────────────────
\usepackage[utf8]{inputenc}
\usepackage[T1]{fontenc}
\usepackage{geometry}
\geometry{
  top=0.9cm,
  bottom=0.9cm,
  left=1.4cm,
  right=1.4cm
}
\usepackage{hyperref}
\hypersetup{
  colorlinks=true,
  urlcolor=blue,
  linkcolor=blue,
  pdftitle={LECA Resurrection Protocol --- One Page Summary},
  pdfauthor={Eric Robert Lawson / OrganismCore}
}
\usepackage{microtype}
\usepackage{parskip}
\usepackage{booktabs}
\usepackage{array}
\usepackage{tabularx}
\usepackage{xcolor}
\usepackage{mdframed}
\usepackage{multicol}
\usepackage{enumitem}
\usepackage{titlesec}
\usepackage{lmodern}

% ── COLOURS ───────────────────────────────────────────────────
\definecolor{headerblue}{RGB}{20,60,140}
\definecolor{statementbox}{RGB}{20,60,140}
\definecolor{rulecolor}{RGB}{20,60,140}

% ── SECTION FORMAT ────────────────────────────────────────────
\titleformat{\section}
  {\small\bfseries\color{headerblue}}
  {}{0em}{}[\vspace{-0.25em}
  \textcolor{rulecolor}{\rule{\linewidth}{0.4pt}}]
\titlespacing{\section}{0pt}{0.35em}{0.15em}

\setlength{\parskip}{0.15em}
\setlength{\parindent}{0pt}
\setlength{\columnsep}{0.9em}
\pagestyle{empty}

% ══════════════════════════════════════════════════════════════
\begin{document}

% ── HEADER ────────────────────────────────────────────────────
\begin{center}
  {\large\bfseries\color{headerblue}
  THE LECA RESURRECTION PROTOCOL}\\[0.1em]
  {\small\color{headerblue}
  Recovery of the Last Eukaryotic Common Ancestor
  from Living Developmental Programs --- One-Page Summary}\\[0.1em]
  {\footnotesize
  Eric Robert Lawson\ $\cdot$\
  OrganismCore (Independent Research)\ $\cdot$\
  ORCID:
  \href{https://orcid.org/0009-0002-0414-6544}
  {0009-0002-0414-6544}\ $\cdot$\
  \href{mailto:OrganismCore@proton.me}
  {OrganismCore@proton.me}\ $\cdot$\
  2026-03-12}
\end{center}

\vspace{0.1em}

% ── THE CLAIM ─────────────────────────────────────────────────
\begin{mdframed}[
  backgroundcolor=statementbox,
  linecolor=statementbox,
  linewidth=0pt,
  innertopmargin=0.3em,
  innerbottommargin=0.3em,
  innerleftmargin=0.7em,
  innerrightmargin=0.7em
]
\begin{center}
\footnotesize\bfseries\color{white}
Every eukaryotic organism carries the Last Eukaryotic Common
Ancestor (\textasciitilde{}1.5\,Ga) as the base state of its
developmental program. Placing any eukaryote in conditions
geometrically derived from that attractor --- before the
experiment runs --- stabilises it there.
This is deliberate de-extinction of the oldest organism on Earth
and the formal definition of a \textbf{universal structural
invariant} observable in every reproductive cycle of every
eukaryote alive --- confirmed by a \textbf{\$200--400}
experiment at any BSL-1 community biology laboratory
in \textbf{days}. Pre-registered 2026-03-12.
\end{center}
\end{mdframed}

\vspace{0.2em}

% ── TWO COLUMNS ───────────────────────────────────────────────
\begin{multicols}{2}

% ── LEFT COLUMN ───────────────────────────────────────────────

\section{The Theorem}

The developmental trajectory from zygote to
adult IS the evolutionary trajectory of the
lineage, compressed into developmental time
(the Reproduction Theorem). Therefore:

\smallskip
{\small
\begin{tabular}{p{0.55cm}p{5.5cm}}
\textbf{E} &
\textbf{Evolutionary grade} --- the attractor
the lineage occupied at that point in its
history. \\[0.2em]
\textbf{D} &
\textbf{Developmental position} --- the
equivalent stage in the developmental
program today. \\
\end{tabular}
}

\smallskip
{\small
$E \equiv D$ (compressed). Arresting at
position $D$ stabilises grade $E$.
The LECA sits at $D{=}0$: before the first
multicellular step, in every eukaryote alive.
\textbf{Conditions derived before the experiment.
Prediction locked before the experiment.
Result confirms the geometry.}
}

{\footnotesize
Pre-registration DOI:\
\href{https://doi.org/10.5281/zenodo.18986790}
{\texttt{10.5281/zenodo.18986790}}
}

\section{The Three Pre-Registered DOIs}

{\footnotesize
\begin{tabular}{p{3.6cm}p{2.4cm}}
\toprule
\textbf{Document} & \textbf{DOI} \\
\midrule
Pre-Registration v1.0 &
\href{https://doi.org/10.5281/zenodo.18986790}
{\texttt{...18986790}} \\[0.1em]
Registered Amendment A1 &
\href{https://doi.org/10.5281/zenodo.18989573}
{\texttt{...18989573}} \\[0.1em]
The Ribosome Is the LUCA &
\href{https://doi.org/10.5281/zenodo.18988844}
{\texttt{...18988844}} \\
\bottomrule
\end{tabular}

\smallskip
Repository:\
\href{https://github.com/Eric-Robert-Lawson/attractor-oncology}
{\texttt{github.com/Eric-Robert-Lawson/}
\texttt{attractor-oncology}}
}

\section{The Five Pre-Registered Predictions}

{\small
\begin{tabular}{p{0.35cm}p{5.7cm}}
\textbf{P1} &
Arrest of any eukaryote before multicellular
organisation produces a single-cell, nucleated,
viable LECA-grade organism. Arm~A alone. \\[0.15em]
\textbf{P2} &
Arrested organisms from animal, plant, and fungal
starting material are \textit{morphologically
indistinguishable} under light microscopy. \\[0.15em]
\textbf{P3} &
\textit{Genomically distinct} by sequencing ---
same attractor, different substrates. \\[0.15em]
\textbf{P4} &
LECA-grade conserved genes active; all
lineage-specific post-LECA genes silenced. \\[0.15em]
\textbf{P5} &
Blind panel cannot identify starting kingdom
from morphology alone (above 1/3 chance). \\
\end{tabular}
}

\columnbreak

% ── RIGHT COLUMN ──────────────────────────────────────────────

\section{What Arm A Alone Establishes}

{\small
Arm A is not a dormancy experiment.
The arrest medium is derived geometrically
from the LECA attractor before the experiment:
Precambrian ionic composition, no nitrogen,
microaerophilic ($<$3\% O$_2$, pre-GOE),
20--22\,\textdegree C.

\textbf{Arm A confirmed = two simultaneous results:}

\textbf{1. De-extinction.}
First deliberate, pre-specified recovery of the
LECA as an intentional target.
\$225\,M has not produced a living mammoth
(extinct 4,000\,yr). Arm~A recovers a 1.5\,Ga
organism for \$200--400 in days --- the oldest
target is the easiest to recover.

\textbf{2. Universal structural invariant defined.}
Developmental arrest is the instrument that
formally defines a structure already present
and observable in every reproductive cycle
of every eukaryote on Earth --- without
intervention. Every developmental dataset
ever collected is retroactively a measurement
of this invariant. Every future developmental
study confirms it. The LUCA runs continuously
in every mitochondrion inside every LECA-grade
cell right now. Once the base state is formally
defined, the invariant is visible everywhere.
}

\section{The Three Experimental Arms}

{\small
\begin{tabular}{p{1.1cm}p{2.05cm}p{1.0cm}p{1.35cm}}
\toprule
\textbf{Arm} & \textbf{Start} &
\textbf{Cost} & \textbf{Timeline} \\
\midrule
A --- Fungal &
\textit{S.~cerevisiae} &
\$200--350 & 24--72\,h \\[0.1em]
B --- Animal &
Mouse zygote &
\$400--600 & 24--72\,h \\[0.1em]
C --- Plant &
\textit{Arabidopsis} &
\$200--400 & 48--72\,h \\
\midrule
\multicolumn{4}{p{5.5cm}}{%
\footnotesize
Arm~A: P1 confirmed. Invariant base state defined.
BSL-1. No credentials required.
Arms A+B+C: morphologically identical across kingdoms,
genomically distinct. Invariant confirmed across
maximum phylogenetic distance.}\\
\bottomrule
\end{tabular}
}

\section{Arm A --- What You Actually Do}

{\small
\textbf{Day~0--3:} Sporulate \textit{S.~cerevisiae}
in 1\% KAc, 25\,\textdegree C, 48--72\,h.
Confirm asci by microscopy.

\textbf{Day~3:} Wash into arrest medium:
10\,mM NaCl, 5\,mM KCl, 2\,mM MgCl$_2$,
1\,mM CaCl$_2$, 0.1\,mM FeSO$_4$,
1\,mM KH$_2$PO$_4$, 0.1\% glucose,
\textbf{no nitrogen}, pH\,6.5--7.0,
20--22\,\textdegree C, $<$3\% O$_2$.

\textbf{Days~4--10:} Observe 24\,h, 48\,h,
72\,h. Phase contrast (40$\times$) + DAPI.
Pass: ${\geq}$80\% single-cell, nucleus
confirmed, ${\leq}$10\% germination,
viable $=$ \textbf{P1 confirmed.}

\textbf{Key control:} At 72\,h add nitrogen
to half the culture --- germination resumes,
proving stabilisation not cell death.

\textbf{Equipment:} Microscope, incubator,
laminar flow. All at ChiTownBio, Chicago.
}

\end{multicols}

\vspace{0.05em}
\noindent\textcolor{rulecolor}
{\rule{\linewidth}{0.4pt}}

% ── FOOTER ────────────────────────────────────────────────────
\begin{center}
\footnotesize
\textit{The geometry came first.\ \
The invariant was always running.\ \
The experiment is the first deliberate observation.\ \
The LECA was never lost. It was never not there.}\\[0.15em]
\href{https://github.com/Eric-Robert-Lawson/attractor-oncology}
{\texttt{github.com/Eric-Robert-Lawson/attractor-oncology}}
\ $\cdot$\
\href{https://doi.org/10.5281/zenodo.18986790}
{\texttt{doi.org/10.5281/zenodo.18986790}}
\ $\cdot$\
\href{https://orcid.org/0009-0002-0414-6544}
{\texttt{ORCID: 0009-0002-0414-6544}}
\end{center}

\end{document}
