\documentclass[9pt, a4paper]{article}

% ── PACKAGES ──────────────────────────────────────────────────
\usepackage[utf8]{inputenc}
\usepackage[T1]{fontenc}
\usepackage{geometry}
\geometry{top=1.4cm, bottom=1.4cm,
          left=1.6cm, right=1.6cm}
\usepackage{hyperref}
\hypersetup{
    colorlinks=true,
    linkcolor=blue,
    urlcolor=blue,
    citecolor=blue,
    pdftitle={FOXA1/EZH2 IHC Classifier ---
              One-Page Summary},
    pdfauthor={Eric Robert Lawson / OrganismCore}
}
\usepackage{booktabs}
\usepackage{array}
\usepackage{parskip}
\usepackage{microtype}
\usepackage{titlesec}
\usepackage{xcolor}
\usepackage{enumitem}
\usepackage{mdframed}
\usepackage{multicol}
\usepackage{fancyhdr}
\usepackage{tabularx}

% ── COLOURS ───────────────────────────────────────────────────
\definecolor{headerblue}{RGB}{20,60,140}
\definecolor{derivebox}{RGB}{235,245,255}
\definecolor{warningbox}{RGB}{255,245,200}
\definecolor{statsbox}{RGB}{248,240,255}
\definecolor{askbox}{RGB}{240,255,240}
\definecolor{equitybox}{RGB}{255,245,230}
\definecolor{mechbox}{RGB}{245,245,255}

% ── PAGE STYLE ────────────────────────────────────────────────
\pagestyle{fancy}
\fancyhf{}
\lhead{\scriptsize\textbf{FOXA1/EZH2 Dual IHC
       Breast Cancer Classifier}}
\rhead{\scriptsize OrganismCore \textbar{}
       Eric Robert Lawson \textbar{} 2026-03-06}
\rfoot{\scriptsize
       \href{mailto:OrganismCore@proton.me}
       {OrganismCore@proton.me} \textbar{}
       ORCID: 0009-0002-0414-6544 \textbar{}
       \href{https://github.com/Eric-Robert-Lawson/attractor-oncology}
       {github.com/Eric-Robert-Lawson/attractor-oncology}}
\renewcommand{\headrulewidth}{0.4pt}
\renewcommand{\footrulewidth}{0.4pt}
\setlength{\headheight}{11pt}
\setlength{\footskip}{14pt}

% ── SECTION FORMATTING ────────────────────────────────────────
\titleformat{\section}
  {\small\bfseries\color{headerblue}}
  {}{0em}{}[\vspace{-4pt}\textcolor{headerblue}
  {\rule{\linewidth}{0.4pt}}]
\titlespacing{\section}{0pt}{4pt}{2pt}

\setlength{\parskip}{2pt}
\setlength{\parindent}{0pt}
\setlength{\columnsep}{0.5cm}

% ══════════════════════════════════════════════════════════════
\begin{document}
\thispagestyle{fancy}

% ── TITLE BLOCK ───────────────────────────────────────────────
\begin{center}
{\large\bfseries\color{headerblue}
FOXA1/EZH2 Dual IHC Breast Cancer Classifier}\\[0.15em]
{\small Two stains. One ratio. Six subtypes.
Same-day results. \$50--\$100 per patient.
TMA-compatible.}
\end{center}

\vspace{0.15em}
\noindent\rule{\textwidth}{1.2pt}
\vspace{0.1em}

% ── DERIVATION BOX — COMPACT ──────────────────────────────────
\begin{mdframed}[backgroundcolor=derivebox,
                 linecolor=blue!50!black,
                 linewidth=0.8pt,
                 innertopmargin=3pt,
                 innerbottommargin=3pt,
                 innerleftmargin=6pt,
                 innerrightmargin=6pt,
                 skipabove=0pt,
                 skipbelow=0pt]
{\small\textbf{\color{headerblue}HOW THIS WAS
DERIVED --- READ THIS FIRST}}

\smallskip
{\small
\textbf{Not found by mining data. Derived from
first principles.}
(1)~Waddington attractor geometry applied to
19,542 single cancer cells identified FOXA1 and
EZH2 as mechanistic opposites on the primary
breast cancer identity axis ---
\textbf{before any clinical dataset was opened.}
(2)~Predicted ordering LumA $>$ LumB $>$ HER2
$>$ TNBC $>$ Claudin-low was
\textbf{locked in a timestamped document before
any confirmatory analysis was run.}
(3)~Tested in seven independent datasets
(${\sim}7{,}500$ patients, 4 platforms):
\textbf{confirmed exactly in all seven.}

\smallskip
\textit{The statistics below are confirmations
of a pre-specified prediction --- not the source
of it. Eliminates overfitting and data dredging
as explanations. Timestamped before-document is
publicly verifiable.}}
\end{mdframed}

\vspace{0.2em}

% ── THREE COLUMN BODY ─────────────────────────────────────────
\begin{multicols}{3}

% ── COL 1: MECHANISM + CLASSIFIES ─────────────────────────────
\section{Biological Mechanism}

{\small
\textbf{FOXA1} --- pioneer transcription factor.
Opens luminal cell identity programme.
High = identity active.

\smallskip
\textbf{EZH2} --- PRC2 catalytic subunit.
Deposits H3K27me3 on FOXA1, GATA3, ESR1
promoters. High = identity silenced.

\smallskip
\textbf{Ratio = which force is winning.}
Determines subtype and treatment vulnerability.

\smallskip
\textit{Independently confirmed: Schade et al.\
(\textit{Nature} 2024); Toska et al.\
(\textit{Nat.\ Med.} 2017). Neither group
knew of this framework.}}

\section{What It Classifies}

{\small
\begin{tabular}{@{}p{1.55cm}p{2.3cm}@{}}
\toprule
\textbf{Subtype} & \textbf{Treatment} \\
\midrule
Luminal A    & CDK4/6i $+$ ET \\[0.15em]
Luminal B    & HDACi $+$ ET \\[0.15em]
HER2-enr.    & Anti-HER2 first \\[0.15em]
TNBC         & Tazemetostat\\
             & $\rightarrow$ fulvestrant \\[0.15em]
Claudin-low  & Anti-TIGIT\\
             & $\rightarrow$ anti-PD-1 \\[0.15em]
ILC          & Fulvestrant $>$ AI \\
\bottomrule
\end{tabular}}

\columnbreak

% ── COL 2: CONFIRMATIONS ──────────────────────────────────────
\section{Confirmations of the\\Pre-Specified Prediction}

\begin{mdframed}[backgroundcolor=statsbox,
                 linecolor=purple!40!black,
                 linewidth=0.6pt,
                 innertopmargin=2pt,
                 innerbottommargin=2pt,
                 innerleftmargin=4pt,
                 innerrightmargin=4pt,
                 skipabove=0pt,
                 skipbelow=2pt]
{\small
\begin{itemize}[noitemsep,
                leftmargin=*,
                topsep=0pt,
                itemsep=1pt]
    \item Ordering: TCGA $n{=}837$,\\
    $p{=}2.87{\times}10^{-103}$
    \item LumA/LumB: METABRIC $n{=}1{,}980$,\\
    $p{=}8.47{\times}10^{-67}$
    \item Survival: 4 cohorts,
    ${\sim}7{,}500$ patients
    \item AUC LumA vs Basal:
    $0.828$--$0.901$
    \item AUC LumA vs LumB:
    $0.796$--$0.873$\\
    {\scriptsize(replicated METABRIC + TCGA
    independently)}
    \item Protein: CPTAC MS $n{=}122$,\\
    $r{=}{-}0.492$, $p{<}0.0001$
    \item Chemo: GSE25066 $n{=}508$,\\
    $p{<}0.0001$
    \item ET: METABRIC $n{=}1{,}104$,\\
    $p{=}0.0027$,
    $\Delta{=}18.7$ months RFS
    \item \textbf{Contradictions: 0}
\end{itemize}}
\end{mdframed}

\section{Not Yet Established}

\begin{mdframed}[backgroundcolor=warningbox,
                 linecolor=orange!60!black,
                 linewidth=0.6pt,
                 innertopmargin=2pt,
                 innerbottommargin=2pt,
                 innerleftmargin=4pt,
                 innerrightmargin=4pt,
                 skipabove=0pt,
                 skipbelow=0pt]
{\small\textbf{IHC H-score cut-points
do not yet exist.}

RNA AUC values confirm the signal.
Thresholds must be derived from stained
tissue vs PAM50 using Youden J.

\textbf{Instrument calibration only.
Same step as ER, PR, HER2, Ki67.
The only remaining step.}}
\end{mdframed}

\columnbreak

% ── COL 3: ASK + WHY IT MATTERS ───────────────────────────────
\section{The Ask}

\begin{mdframed}[backgroundcolor=askbox,
                 linecolor=green!50!black,
                 linewidth=0.6pt,
                 innertopmargin=2pt,
                 innerbottommargin=2pt,
                 innerleftmargin=4pt,
                 innerrightmargin=4pt,
                 skipabove=0pt,
                 skipbelow=2pt]
{\small
$300$--$400$ archived FFPE cases with
PAM50 (whole section \textbf{or} TMA).

FOXA1 CST D4E2 \#53528 $+$
EZH2 CST D2H1 \#5246.

H-score blinded to subtype.

No new patients.
No treatment changes.

\textbf{\$5,000--\$15,000 consumables.
3--6 months.}

\smallskip
\textit{In return: Protocol v3.0,
calibration study design and analysis framework,
co-authorship, open access (MIT).}}
\end{mdframed}

\section{Why This Matters}

\begin{mdframed}[backgroundcolor=equitybox,
                 linecolor=orange!50!black,
                 linewidth=0.6pt,
                 innertopmargin=2pt,
                 innerbottommargin=2pt,
                 innerleftmargin=4pt,
                 innerrightmargin=4pt,
                 skipabove=0pt,
                 skipbelow=2pt]
{\small
PAM50: \$3,000--\$4,000.
Proprietary RNA platform.
Inaccessible to the majority
of the world's breast cancer patients.

\smallskip
This test: \$50--\$100.
Standard IHC. Any laboratory.
Same day. No RNA. No shipping.

\smallskip
\textbf{${\sim}1.2$--$1.5$M patients/year
receive no molecular subtype information.
This test, on validation,
reaches all of them.}}
\end{mdframed}

\section{Published Evidence}

{\small
\textbf{CS-LIT-1}
(${\sim}7{,}500$ patients):\\
\href{https://doi.org/10.5281/zenodo.18883922}
{\scriptsize doi.org/10.5281/zenodo.18883922}

\smallskip
\textbf{CS-LIT-22} (IHC tool):\\
\href{https://doi.org/10.5281/zenodo.18892788}
{\scriptsize doi.org/10.5281/zenodo.18892788}

\smallskip
\textbf{Full repository}:\\
\href{https://github.com/Eric-Robert-Lawson/attractor-oncology}
{\scriptsize github.com/Eric-Robert-Lawson/\\
attractor-oncology}}

\end{multicols}

\end{document}
